### 2012-02-24 Krantz-v3/Chapters/chapter1/discussion.tex
Added booke chapter.

The SLDA model family, of which HSLDA is a member, can be understood in two
different ways. One way is to see it as a family of topic models that improve
on the topic modeling performance of LDA via the inclusion of observed
supervision. An alternative, complementary way is to see it as a set of models
that can predict labels for bag-of-word data. A large diversity of problems
can be expressed as label prediction problems for bag-of-word data. A
surprisingly large amount of that kind of data possess structured labels,
either hierarchically constrained or otherwise. That HSLDA directly addresses
this kind of data is a large part of the motivation for this work. That it
outperforms more straightforward approaches should be of interest to
practitioners.

Variational Bayes has been the predominant estimation approach applied to sLDA
models. Hierarchical probit regression makes for tractable Markov chain Monte
Carlo SLDA inference, a benefit that should extend to other sLDA models should
probit regression be used for response variable prediction there too.

The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
it is better to do full joint estimation of HSLDA.  An alternative
interpretation of the same results is that, if one is more sensitive to the
performance gains that result from exploiting the structure of the labels, then
one can, in an engineering sense, get nearly as much gain in label prediction
performance by first fitting LDA and then fitting a hierarchical probit
regression.  There are applied settings in which this could be advantageous.

Extensions to this work include unbounded topic cardinality variants and
relaxations to different kinds of label structure.  Unbounded topic cardinality
variants pose interesting inference challenges.  Utilizing different kinds of
label structure is possible within this framework, but requires relaxing some
of the simplifications we made in this paper for expositional purposes.

>>>>

The SLDA model family, of which HSLDA is a member, can be understood in two
different ways. One way is to see it as a family of topic models that improve
on the topic modeling performance of LDA via the inclusion of observed
supervision. An alternative, complementary way is to see it as a set of models
that can predict labels for bag-of-word data. A large diversity of problems
can be expressed as label prediction problems for bag-of-word data. A
surprisingly large amount of that kind of data possess structured labels,
either hierarchically constrained or otherwise.  HSLDA directly addresses
this kind of data and works well in practice. That it
outperforms more straightforward approaches should be of interest to
practitioners.
%
%Variational Bayes has been the predominant estimation approach applied to sLDA
%models. Hierarchical probit regression makes for tractable Markov chain Monte
%Carlo SLDA inference, a benefit that should extend to other sLDA models should
%probit regression be used for response variable prediction there too.
%
%The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
%it is better to do full joint estimation of HSLDA.  An alternative
%interpretation of the same results is that, if one is more sensitive to the
%performance gains that result from exploiting the structure of the labels, then
%one can, in an engineering sense, get nearly as much gain in label prediction
%performance by first fitting LDA and then fitting a hierarchical probit
%regression.  There are applied settings in which this could be advantageous.

Extensions to this work include unbounded topic cardinality variants and
relaxations to different kinds of label structure.  Unbounded topic cardinality
variants pose interesting inference challenges.  Imposing different kinds of
label structure constraints is possible within this framework, but requires relaxing some
of the assumptions we made in deriving the sampling distributions for HSLDA inference.

### 2012-02-24 Krantz-v3/Chapters/chapter1/experiments.tex
Added booke chapter.

We applied HSLDA to data from two domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

@@

\subsection{Data and Pre-Processing}

\subsubsection{Discharge Summaries and ICD-9 Codes}

Discharge summaries are authored by clinicians to summarize patient
hospitalization courses. The summaries typically contain a record of patient
complaints, findings and diagnoses, along with treatment and hospital course.
For each hospitalization, trained medical coders review the information in the
discharge summary and assign a series of diagnoses codes. Coding follows the
ICD-9-CM controlled terminology, an international diagnostic classification for
epidemiological, health management, and clinical
purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  The ICD-9 codes are organized in a rooted-tree structure, with each
edge representing an is-a relationship between parent and child, such that the
parent diagnosis subsumes the child diagnosis. For example, the code for
{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
pneumonia,'' where the former is a type of the latter.  It is worth noting that
the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
tend to be more specific than sensitive in their
assignments~\cite{Birmetal2005}, and sometimes make
mistakes~\cite{FarzandipourEtAl10}.

The task of automatic ICD-9 coding has been investigated in the clinical
domain.  Methods range from manual
rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
Other work had leveraged larger datasets and experimented with K-nearest
neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
well as simple keyword mappings, all with promising
results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Lita2008}.%Ruch2008 omitted

@@

Our dataset was gathered from the NewYork-Presbyterian Hospital clinical data warehouse. 
It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. All included discharge summaries had associated ICD-9 Codes.
Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

@@

%For each hospitalization there are usually several ICD-9 codes assigned
%for billing purposes. These codes are known to be quite specific but
%not very sensitive \citep{}. Regardless of that fact,
%this is one of the only sources for information on patient diagnoses
%aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

\subsubsection{Product Descriptions and Categorizations}

Amazon.com, an online retail store, organizes its catalog of products in a
mulitply-rooted hierarchy and provides textual product descriptions 
for most products. Products can be discovered by users
through free-text search and product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

In this experiment, we obtained Amazon.com product categorization data from 
the Stanford Network Analysis Platform (SNAP) dataset~\citep{SNAP}. 
Product descriptions were obtained separately from the
Amazon.com website directly. We limited our dataset to the collection of DVDs
in the product catalog.

%We were able to deduce the structure of the
%hierarchy for the Amazon.com products because all ancestors in the
%hierarchy were included with each category label. For example, {}``DVD / Genres
%/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
%for the DVD, {}``The Time Machine.''

Our dataset contains 15,130 product descriptions for training and 1,000
for testing. The product descriptions are shorter than the discharge summaries
(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique categories.
Products are assigned on average 9.01 categories (std dev=4.91). The vocabulary
consists of the most frequent 30,000 words omitting stopwords.

@@

We evaluated HSLDA along with two closely related models against
the two datasets. The comparison models included sLDA with independent
regressors (hierarchical constraints on labels ignored) and HSLDA fit by first
performing LDA then fitting tree-conditional regressions. These models were
chosen to  highlight several aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

sLDA with independent regressors is the most salient comparison model
for our work. The distinguishing factor between HSLDA and sLDA is the
additional structure imposed on the label space, a distinction that we
hypothesized would result in a difference in predictive performance.

There are two components to HSLDA, LDA and a hierarchically constrained
response. The second comparison model is HSLDA fit by performing LDA first
followed by performing inference over the hierarchically constrained label
space. In this comparison model, the separate inference processes do not allow
the responses to influence the low dimensional structure inferred by LDA.
Combined inference has been shown to improve performance in sLDA
\citep{BleiMcAuliffe2008}. This comparison model examines not the structuring
of the label space, but the benefit of combined inference over both the
documents and the label space.

@@

For all three models, particular attention was given to the settings of the 
prior parameters for the regression coefficients. These parameters implement an
important form of regularization in HSLDA. In the setting where there are no
negative labels, a Gaussian prior over the regression parameters with a
negative mean implements a prior belief that missing labels are likely to be
negative. Thus, we evaluated model performance for all three models with a
range of values for $\mu$, the mean prior parameter for regression coefficients 
($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$).

The number of topics for all models was set to 50, the prior distributions of
$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
$p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1 and a
scale parameters of 1000.

\begin{figure}[h]
\begin{center}
%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][Clinical data performance.]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][Retail product performance.]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
%\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
\caption{ROC curves for out-of-sample label prediction varying $\mu$, the prior mean of the regression parameters. 
In both figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical 
constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
tree-conditional regressions.}
%\label{fig:main_results}
\end{center}
\end{figure}

@@

We evaluated our model, HSLDA, against the comparison models with a focus on
predictive performance on held-out data. Prediction performance was measured
with standard metrics -- sensitivity (true positive rate) and 1-specificity
(false positive rate).

The gold standard for comparison was derived from the testing set in each
dataset. To make the comparison as fair as possible among models, ancestors of
observed nodes in the label hierarchy were ignored, observed nodes were
considered positive and descendents of observed nodes were considered to be
negative. Note that this is different from our treatment of the observations during inference.
Since the sLDA model does not enforce the hierarchical
constraints, we establish a more equal footing by
considering only the observed labels as being positive, despite the fact
that, following the hierarchical constraints, ancestors must also be positive.
Such a gold standard will likely inflate the number of false positives because
the labels applied to any particular document are usually not as complete as
they could be.  ICD-9 codes, for instance, lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as
false positives~\citep{Birmetal2005}.
However, given that the label space is often large (as in our examples) it is a
moderate assumption that erroneous false positives should not skew results
significantly.

Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
for each test document, $\hat{d}$. For efficiency, the expectation of this
probability distribution was estimated in the following way. Expectations 
of $\mathbf{\bar{z}}_{\hat{d}}$ and $\boldsymbol{\eta}_l$ were estimated with samples
from the posterior. Using these expectations, we performed Gibbs sampling over
the hierarchy to acquire predictive samples for the documents in the test set.
The true positive rate was calculated as the average expected labeling for
gold standard positive labels. The false positive rate was calculated 
as the average expected labeling for gold standard negative labels.

@@

\begin{figure}[t]
%tbp] %  figure placement: here, top, bottom, or page
\centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC curve for out-of-sample ICD-9 code prediction varying auxiliary variable threshold. $\mu = -1.0$ for all three models in this figure.}
\label{fig:results_auxiliary_var} 
\end{figure}

Figure~\ref{fig:1a} demonstrates the performance of the model on the clinical data as an ROC curve varying $\mu$.  For instance, 
a hyperparameter setting of $\mu=-1.6$ yields the following performance:
the full HSLDA model had a 
true positive rate of 0.57 and a false positive rate of 0.13, the sLDA model
had a true positive rate of 0.42 and a false positive rate of 0.07, and the
HSLDA model where LDA and the regressions were fit separately had a true
positive rate of 0.39 and a false positive rate of 0.08. These points are highlighted in Figure~\ref{fig:1a}.

@@

Figure~\ref{fig:1b} demonstrates the performance of the model on the retail product data as an ROC curve also varying $\mu$. For
instance, a hyperparameter setting of $\mu=-2.2$ yields the following
performance: the full HSLDA model had a true positive rate of 0.85 
and a false positive rate of 0.30, the sLDA model had a true positive
rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.77
and a false positive rate of 0.16. These results follow a similar pattern to the clinical data. These points are highlighted in Figure~\ref{fig:1b}.

Figure \ref{fig:results_auxiliary_var} shows the predictive performance of HSLDA 
relative to the two comparison models on the clinical dataset as a function of the auxiliary variable threshold. 
For low values of the auxiliary variable threshold, the models predict labels
in a more sensitive and less specific manner, creating the points in the upper
right corner of the ROC curve. As the auxiliary variable threshold is
increased, the models predict in a less sensitive and more specific manner,
creating the points in the lower left hand corner of the ROC curve. HSLDA with full joint inference outperforms sLDA with
independent regressors as well as HSLDA with separately trained
regression.

>>>>

Having described HSLDA and given details on how to do inference in it, we turn to demonstrations of using HSLDA to solve problems from  two different domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

@@

\subsection{Hospital Discharge Summaries and ICD-9 Codes}

 Despite the growing emphasis on meaningful
use of technology in medicine, many aspects of medical record-keeping
remain a manual process. In the US, diagnostic coding for billing and insurance
purposes is often handled by professional medical coders who must
explore a patient's extensive clinical record before assigning the
proper codes. %So while electronic health records (EHRs) should be
%adopted by most medical institutions within the next several years,
%largely due to the provisions of HITECH under the American Recovery
%and Reinvestment Act \cite{Blumenthal2009}, there has been little
%movement forward in automating medical coding.

%Similarly, viewing the distribution of topics over discharge summaries
%may reveal information about the latent structure of clinician documentation.
%Lastly, the SLDA model would provide a novel approach to dealing with
%the problem of high dimensionality when representing narrative text
%in a vector space specifically by reducing dimensions from an entire
%vocabulary of potentially tens of thousands of words to a set of several
%dozen topics.

A specific example of this involves labeling of hospital discharge summaries.  These summaries are authored by clinicians to summarize patient
hospitalization courses. They typically contain a record of patient
complaints, findings and diagnoses, along with treatment and hospital course.  The kind of text one might expect to find in such a discharge summary is illustrated by this made-up snippet
\begin{quote}
{ADLER Dischage summary text description snippet }
\end{quote} 
For each hospitalization, trained medical coders review the information in the
discharge summary and assign a series of diagnoses codes. Coding follows the
ICD-9-CM controlled terminology, an international diagnostic classification for
epidemiological, health management, and clinical
purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  These ICD-9 codes are organized in a rooted-tree structure, with each
edge representing an is-a relationship between parent and child, such that the
parent diagnosis subsumes the child diagnosis. For example, the code for
{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
pneumonia,'' where the former is a type of the latter.  A representative sub-tree of the ICD-9 code tree is shown in Figure ADLER\ref{ADLER}. It is worth noting that
the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
tend to be more specific than sensitive in their
assignments~\cite{Birmetal2005}, and sometimes make
mistakes~\cite{FarzandipourEtAl10}.

An automated process would ideally produce a more complete and accurate
diagnosis lists. %Also, this model will reveal information about the
%medical records themselves. For example, we may gain an understanding
%of what a specific code actually means in terms of clinical narratives.
The task of automatic ICD-9 coding has been investigated in the clinical
domain.  Methods used to solve this problem (besides HSLDA) range from applying manually derived coding rules
rules to applications of online rule learning approaches \cite{Crammer2007,Goldstein2007,Farkas2008}.
Many classification schemes have been applied to this problem: K-nearest
neighbor , Naive Bayes, support vector machines, Bayesian Ridge Regression, as
well as simple keyword mappings, all with promising
results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Lita2008}.%Ruch2008 omitted

@@

The specific dataset we report results for in this chapter was gathered from the NewYork-Presbyterian Hospital clinical data warehouse. 
It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. All included discharge summaries had associated ICD-9 Codes.
Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

@@

Here HSLDA is evaluated as a way to understand and model the relationship between a discharge summary and the ICD-9 codes that should be assigned to it.  We show promising results for automatically assigning ICD-9 codes to hospital discharge records.

%For each hospitalization there are usually several ICD-9 codes assigned
%for billing purposes. These codes are known to be quite specific but
%not very sensitive \cite{}. Regardless of that fact,
%this is one of the only sources for information on patient diagnoses
%aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

\subsection{Product Descriptions and Catalogs}

Many web-retails store and organize their catalog of products in a
mulitply-rooted hierarchy in addition to providing textual product descriptions 
for most products. Products can be discovered by users
through free-text search and product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

Amazon.com is one such retailer.  Their product categorization data is available as part of the 
 Stanford Network Analysis Platform (SNAP) dataset~\cite{SNAP}.    A representative sub-tree of the amazon.com DVD produce categorization tree is shown in Figure ADLER\ref{ADLER}.  
Product descriptions were obtained separately from the
Amazon.com website directly. Once such description is
\begin{quote}
{ADLER DVD description }
\end{quote}
We study the collection of DVDs
in the product catalog specifically.
%We were able to deduce the structure of the
%hierarchy for the Amazon.com products because all ancestors in the
%hierarchy were included with each category label. For example, {}``DVD / Genres
%/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
%for the DVD, {}``The Time Machine.''
The resulting dataset contains 15,130 product descriptions for training and 1,000
for testing. The product descriptions consist of
91.89 terms on average (std dev=53.08). Overall, there are 2,691 unique categories.
Products are assigned on average 9.01 categories (std dev=4.91). The vocabulary
consists of the most frequent 30,000 words omitting stopwords.

HSLDA is used here to understand and model the relationship between the product text description and the products' positioning in the product hierarchy.  We show how to automatically situate a product in a hierarchal product catalog.

@@

We compare HSLDA to two closely related models. The comparison models are SLDA with independent
regressors (hierarchical constraints on labels ignored,  i.e. the regression is not conditional) and HSLDA fit by first
performing LDA then fitting tree-conditional regressions (rather than jointly inferring the topics and the regression coefficients). These models were
chosen because they are the strongest available competitors and because they  highlight several pedagogical aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

SLDA with independent regressors is the most salient comparison model
for our work. The distinguishing factor between HSLDA and SLDA is the
additional structure imposed on the label space, a distinction that in developing HDSLA we
hypothesized would result in a difference in predictive performance.

 The second comparison model, HSLDA fit by performing LDA first
followed by performing inference over the hierarchically constrained label
space, does not allow
the responses to influence the topics inferred by LDA.
Combined inference has been shown to improve performance in SLDA
\cite{BleiMcAuliffe2008}. This comparison model doesn't examine the value of utilizing the structured nature 
of the label space, instead it highlights the benefit of combined inference over both the
documents and the label space.

@@

For all three models, particular attention was given to the settings of the 
prior parameters for the regression coefficients. These parameters implement an
important form of regularization in HSLDA. In the setting where there are no
negative labels, a Gaussian prior over the regression parameters with a
negative mean implements a prior belief that missing labels are likely to be
negative. Thus, we show model performance for all three models with a
range of values for $\mu$, the mean prior parameter for regression coefficients 
($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$).

The number of topics for all models was set to $K=50$, the prior distributions of
$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
$p\left(\gamma\right)$ were all chosen to be gamma with a shape parameter of 1 and a
scale parameter of 1000.

\begin{figure}[h]
\begin{center}
%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][Clinical data performance.]{\label{fig:1a}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/clin_pred_varying_mu}}
%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][Retail product performance.]{\label{fig:1b}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/amazon_pred_varying_mu}}
%\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
\caption{ROC curves for HSLDA out-of-sample label prediction varying $\mu$, the prior mean of the regression parameters. 
In both figures, solid is HSLDA, dashed are independent regressors + SLDA (hierarchical 
constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
tree-conditional regressions.}
%\label{fig:main_results}
\end{center}
\end{figure}

@@

We are particularly interested in
predictive performance on held-out data. Prediction performance was measured
with standard metrics -- sensitivity (true positive rate) and 1-specificity
(false positive rate).

In each case the gold standard for testing was derived from the test data. To make the comparison as antagonistic to HSLDA as possible (relative to the other models), ancestors of
observed nodes in the label hierarchy were ignored, observed nodes were
considered positive and descendants of observed nodes were assumed to be
negative. Note that this is different from our treatment of the observations during inference where we marginalize over possible settings of unobserved labels.
For instance, as the SLDA model does not enforce hierarchical
label constraints, when we consider only observed nodes we penalize HSLDA.  This is because
 the is-a hierarchical constraints say that the ancestors of positively labeled nodes must also be positive which the SLDA model cannot guarantee.
Another antagonism of this gold standard is that it is likely to inflate the number of false positives because
the labels applied to any particular document are usually not as complete as
they could be.  ICD-9 codes, for instance, are known to lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as
false positives~\cite{Birmetal2005}.
However, given that the label space is often large (as in our examples) it is a
reasonable assumption that erroneous false positives should not skew results
significantly.

Predictive performance in HSLDA is evaluated by computing
 \[p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)\] 
for each test document $\hat{d}$ for each observed label $y_{l,\hat{d}}$ (given the test document words). For efficiency, the expectation of this
probability distribution was approximated in the following way. Expectations 
of $\mathbf{\bar{z}}_{\hat{d}}$ and $\boldsymbol{\eta}_l$ were estimated with samples
from the posterior. Fixing these expectations, we performed Gibbs sampling over
the hierarchy to acquire predictive samples for the documents in the test set.
The true positive rate was calculated as the average expected labeling for
gold standard positive labels. The false positive rate was calculated 
as the average expected labeling for gold standard negative labels.

@@

\begin{figure}[t]
%tbp] %  figure placement: here, top, bottom, or page
\centering \includegraphics[scale=0.4]{Chapters/chapter1/figs/ROC_comparison_leafs} \caption{ROC curve for out-of-sample ICD-9 code prediction varying auxiliary variable threshold. $\mu = -1.0$ for all three models in this figure.}
\label{fig:results_auxiliary_var} 
\end{figure}

Figure~\ref{fig:1a} demonstrates the performance of the model on the clinical data as an ROC curve varying $\mu$.  For instance, 
a hyperparameter setting of $\mu=-1.6$ yields the following performance:
the full HSLDA model had a 
true positive rate of 0.57 and a false positive rate of 0.13, the SLDA model
had a true positive rate of 0.42 and a false positive rate of 0.07, and the
HSLDA model where LDA and the regressions were fit separately had a true
positive rate of 0.39 and a false positive rate of 0.08. These points are highlighted in Figure~\ref{fig:1a}.

@@

Example topics (as word lists) learned for the discharge data are given below.  These word lists are computed by ADLER

\begin{extract}
ADLER
\end{extract}

ADLER could you write a sentence about how these are cleaner than the SLDA or LDA topics?

Figure~\ref{fig:1b} demonstrates the performance of the model on the retail product data as an ROC curve also varying $\mu$. For
instance, a hyperparameter setting of $\mu=-2.2$ yields the following
performance: the full HSLDA model had a true positive rate of 0.85 
and a false positive rate of 0.30, the SLDA model had a true positive
rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.77
and a false positive rate of 0.16. These results follow a similar pattern to the clinical data. These points are highlighted in Figure~\ref{fig:1b}.

Example topics (as word lists) learned for the amazon.com data are given below.  T

\begin{extract}
ADLER
\end{extract}

Figure \ref{fig:results_auxiliary_var} shows the predictive performance of HSLDA 
relative to the two comparison models on the clinical dataset as a function of the auxiliary variable threshold. 
For low values of the auxiliary variable threshold, the models predict labels
in a more sensitive and less specific manner, creating the points in the upper
right corner of the ROC curve. As the auxiliary variable threshold is
increased, the models predict in a less sensitive and more specific manner,
creating the points in the lower left hand corner of the ROC curve. HSLDA with full joint inference outperforms SLDA with
independent regressors as well as HSLDA with separately trained
regression.

### 2012-02-24 Krantz-v3/Chapters/chapter1/introduction.tex
Added booke chapter.

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper, we focus on
unstructured text data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes.
In this work we show how to combine 
these two sources of information using a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. The models and techniques that we develop in
this paper are applicable to other data as well, namely, any unstructured
representations of data that have been hierarchically classified (e.g.,~image
catalogs with bag-of-feature image representations).

There are several challenges entailed in incorporating a hierarchy of labels
into the model. Among them, given a large set of potential labels (often thousands), each
instance has only a small number of labels associated to it. Furthermore, there are no
naturally occurring negative labeling in the data, and the absence of a label
cannot always be interpreted as a negative labeling.

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. For the sake of
simplicity, we focus on is-a hierarchies, but the model can be
applied to other structured label spaces. We extend supervised latent Dirichlet
allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
supervision. We propose an efficient way to incorporate hierarchical
information into the model. We hypothesize that the context of labels within
the hierarchy provides valuable information about labeling.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. Our results show that a joint, hierarchical model outperforms a
classification with unstructured labels as well as a disjoint model, where the
topic model and the hierarchical classification are inferred
independently of each other.

The remainder of this paper is as follows. Section~\ref{sec:model}
introduces hierarchically supervised LDA (HSLDA), while
Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
Section~\ref{sec:related_work} reviews related work, and
Section~\ref{sec:experiments} shows results from applying HSLDA to health care
and web retail data.

%In this paper we describe the use of a topic model based on supervised
%latent Dirichlet allocation (sLDA) to identify topics within narrative
%discharge summaries and to automate the assignment of diagnostic codes,
%specifically International Classification of Disease 9th Revision,
%Clinical Modification (ICD-9-CM) codes. There are a number of advantages
%to this approach. First, manually coding diagnoses is a time-consuming
%and notoriously unreliable process. Many diagnoses are omitted in
%the final record, and a high error rate is found even in the principal
%diagnoses \citep{Surjan1999}.

@@

% Informatics journal paper
%
% Despite the growing emphasis on meaningful
%use of technology in medicine, many aspects of medical record-keeping
%remain a manual process. Diagnostic coding for billing and insurance
%purposes is often handled by professional medical coders who must
%explore a patient's extensive clinical record before assigning the
%proper codes. So while electronic health records (EHRs) should be
%adopted by most medical institutions within the next several years,
%largely due to the provisions of HITECH under the American Recovery
%and Reinvestment Act \citep{Blumenthal2009}, there has been little
%movement forward in automating medical coding.

>>>>

%This chapter covers one way to utilize supervision in the form of
%hierarchical and (often) multiple labelings in a similar manner. 
Consider web
retail data. Web retailers often have both a browse-able product hierarchy and
free-text descriptions for all products they sell. The situation of each
product in a product hierarchy (often multiply situated) constitutes a
multiple, hierarchical labeling of the free-text product descriptions.  One might assume that the free-text descriptions of all of the products in a particular node in the product hierarchy are related.  Basketballs are probably described using language that is similar to that used to describe other basketballs, other balls, and more general sporting goods.   In both the lay and technical senses, similar products should have product descriptions that share topics.   What is more, if topic proportions are indicative of the text describing products that are grouped together, it should be possible to use those proportions to decide (via classification) whether or not a particular product should be situated at a particular node in the product hierarchy.  Conversely, that certain groups of products are known to be clustered together should inform the kinds of topics that are inferred from the product descriptions.

%hierarchical labels should, at least in theory, provide
%better supervision than the simpler unstructured labels previously considered.
%Results from applying our model to both medical record and web retail data
%suggests that this is likely the case. In particular, we observe gains in our
%primary goal of out-of-sample label prediction that result specifically from
%leveraging hierarchical supervision.

There are many kinds of problems that have the same characteristics as this: any data  that consists of free text that has been partially or
completely categorized by human editors for instance; more fully
% In this paper we focus on
any bag-of-words data that has been, at least in part, 
categorized. 
 Examples include but are not limited to webpages and curated
hierarchical directories of the same \cite{DMOZ}, product descriptions and
catalogs, (e.g.~\cite{AMAZON} as available from \cite{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes.
The model we cover in this chapter shows one way to combine 
these two sources of information into a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. %The models and techniques that we develop in
%this paper are applicable to other data as well, namely, any unstructured
%representations of data that have been hierarchically classified (e.g.,~image
%catalogs with bag-of-feature representations).

%There are several challenges entailed in incorporating a hierarchy of labels
%into the model. Among them, given a large set of potential labels (often thousands), each
%instance has only a small number of labels associated to it. Furthermore, there are no
%naturally occurring negative labeling in the data, and the absence of a label
%cannot always be interpreted as a negative labeling.

%Our work operates within the framework of topic modeling. Our approach learns
%topic models of the underlying data and labeling strategies in a joint model,
%while leveraging the hierarchical structure of the labels. 
In this chapter we focus on ``is-a'' hierarchically structured label spaces, but label spaces that are structured differently may be accommodated with relatively minor modifications. 
%
%We extend supervised latent Dirichlet
%allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
%supervision. We propose an efficient way to incorporate hierarchical
%information into the model. We hypothesize that the context of labels within
%the hierarchy provides valuable information about labeling. 
%
We show results from modeling real-world datasets in the clinical and web
retail domains. These results provide evidence that the two views (text and labels) mutually benefit  multi-label
classification.%In fact, it will be shown that the hierarchical supervised model outperforms a
That is, modeling the joint  is better than learning 
topic models and hierarchical classifiers
independently.

The remainder of this chapter is arranged as follows. Section~\ref{sec:model}
introduces hierarchically supervised LDA (HSLDA), while
Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
Section~\ref{sec:related_work} reviews related work, and
Section~\ref{sec:experiments} shows results from applying HSLDA to health care
and web retail data.

%In this paper we describe the use of a topic model based on supervised
%latent Dirichlet allocation (sLDA) to identify topics within narrative
%discharge summaries and to automate the assignment of diagnostic codes,
%specifically International Classification of Disease 9th Revision,
%Clinical Modification (ICD-9-CM) codes. There are a number of advantages
%to this approach. First, manually coding diagnoses is a time-consuming
%and notoriously unreliable process. Many diagnoses are omitted in
%the final record, and a high error rate is found even in the principal
%diagnoses \cite{Surjan1999}.

@@

% Informatics journal paper
%
% Despite the growing emphasis on meaningful
%use of technology in medicine, many aspects of medical record-keeping
%remain a manual process. Diagnostic coding for billing and insurance
%purposes is often handled by professional medical coders who must
%explore a patient's extensive clinical record before assigning the
%proper codes. So while electronic health records (EHRs) should be
%adopted by most medical institutions within the next several years,
%largely due to the provisions of HITECH under the American Recovery
%and Reinvestment Act \cite{Blumenthal2009}, there has been little
%movement forward in automating medical coding.
